---
title: "Week 6 Assignment"
author: "ESHA"
date: "2025-03-02"
output: html_document
---

# Install necessary packages
```{r}
install.packages("data.table")
install.packages("geosphere")

library(data.table)
library(geosphere)
install.packages("GA")
library(GA)
install.packages("purrr")
library(purrr)
install.packages("sf")
library(sf)
```
```{r}
library(data.table)
library(geosphere)
library(GA)
library(purrr)
library(sf)
```

```{r}
install.packages("parallel")
install.packages("doParallel")
```

getwd()

setwd("C:/Users/palad/Downloads/Mo_pop_Sim")

setwd("C:/Users/palad/Downloads/MO_2018_Federally_Qualified_Health_Center_Locations/MO_2018_Federally_Qualified_Health_Center_Locations")





```{r}
## Load the population data
pop.data <- data.table::fread("C:/Users/palad/Downloads/Mo_pop_Sim/Mo_pop_Sim.csv")
str(pop.data)
frac <- 0.0001
small.data <- pop.data[sample(1:nrow(pop.data),
                               size = round(nrow(pop.data) * frac),
                               replace = F), ## extract a sample of randomlychosen 1% rows
                        ]  ## and choose all columns

```



```{r}

## Load the FQHC data
data_path <- ("C:/Users/palad/Downloads/MO_2018_Federally_Qualified_Health_Center_Locations/MO_2018_Federally_Qualified_Health_Center_Locations")
fqhc.data <- data.table(as.data.frame(st_read(data_path,
                     'MO_2018_Federally_Qualified_Health_Center_Locations')))
names(fqhc.data)
```




```{r}
fqhc.data <- fqhc.data[, .(Facility, Longitude, Latitude)]
small.data <- small.data[, .(long, lat)]  # Use small population data

# Function to calculate the number of residents within a 30-mile radius (~48,280 meters)
count_residences <- function(fqhc, pop_data) {
  distances <- distHaversine(matrix(c(pop_data$long, pop_data$lat), ncol = 2),
                             matrix(c(fqhc$Longitude, fqhc$Latitude), ncol = 2))
  return(sum(distances <= 48280))  # Count residents within the 30-mile radius
}

# Fitness function for GA: sum of residents covered by selected FQHCs
fitness_function <- function(selected_indices) {
  selected_fqhc <- fqhc.data[selected_indices, ]
  total_coverage <- sum(sapply(1:nrow(selected_fqhc), function(i) count_residences(selected_fqhc[i], small.data)))
  return(total_coverage)
}
```


```{r}
library(parallel)
library(doParallel)
```


```{r}
library(GA)
library(sf)
library(data.table)

```

```{r}
# Run Genetic Algorithm
ga_result <- ga(type = "permutation",
                fitness = function(indices) fitness_function(indices),
                lower = 1, upper = nrow(fqhc.data), 
                popSize = 50, maxiter = 100, run = 10,
                parallel = TRUE, pmutation = 0.2,
                names = fqhc.data$Facility)

# Get the best 8 selected locations
best_indices <- ga_result@solution[1, ]
best_fqhc <- fqhc.data[best_indices, ]

# Display selected FQHC locations

print(best_fqhc)

```

  
For selecting 8 optimal FQHCs for specialized mental health services, we use population density within a 30-mile radius as the fitness criterion. This approach ensures that services are placed where they can benefit the most people.  

Key Assumptions:
Higher population density indicates higher service demand  
Residents are more likely to use services within 30 miles.  
The need for mental health services is evenly distributed across the population.  
Population distribution remains static during analysis.  

Fitness Function : 
Metric: The total number of residents within a 30-mile radius** of each FQHC.  
Goal: Maximize this number to select the 8 FQHCs serving the highest number of residents.  












